Propositions as Sessions, Mechanically

### 2015-04-11 paper/formalisation.tex
Described Coq version of ccoutput (added its forgotten figure).

\textbf{Weakening and Contraction.} Initially, an implicit notion of weakening
was chosen so as to be in line with \citeauthor{Wadler:2014}'s
definitions. However, explicit weakening is required to differentiate the
principal cut elimination rules for the interactions between servers and
clients; spawning and garbage collection (weakening). The paper presentation
does not specify where weakening could occur; it is tacitly assumed weakening
can always be pushed further up the derivation tree to a base rule
e.g. \coqe$cp_fwd$ rule.

@@

\textbf{Empty Choice.} The empty choice rule may only have a single binding in
the context (cf. an arbitrary context in \citeauthor{Wadler:2014}'s
presentation~\cite{Wadler:2014}). Adding an arbitrary environment to the empty
choice constructor would prohibit \coqe$in_env_fv$ since its process no longer
contains all the names of its environment. Not only would weakening be
inadmissible if \coqe$in_env_fv$ did not hold, but also axiom cut reduction:
\begin{coq}
Lemma reduce_axcut:
  forall P A (w : atom) $\Gamma$
         (NFV: w $\notin$ fv_proc P)
         (WT: $\nu$ A.(w $\link$ 0 $\mid$ P) $\tpcp$ $\Gamma$),
    (open_proc P w) $\tpcp$ $\Gamma$.
\end{coq}

@@

@@

Regarding commuting conversions, many require analysing permutations
containing singleton bindings. Consider the following commuting conversion:
\begin{coq}
  Lemma reduce_cc_input:
    forall P Q (x:atom) A B $\Gamma$
           (WT: $\nu$ A.(x(B).P $\mid$ Q) $\tpcp$ $\Gamma$),
      x(B).($\nu$ A.($\open{0}{1}{P}$ $\mid$ Q)) $\tpcp$ $\Gamma$.
\end{coq}

Afer inverting \key{WT} and specialising the resulting hypotheses, regarding
the subprocesses, with fresh variables the proof state is as follows:

\begin{tabular}{l}
\begin{coq}
1 subgoal
(* ... other elements of the proof context ... *)
CPQ : open_proc Q y $\tpcp$ y$\sim$A++$\Delta$Q
CPP0 : $\open{0}{z}{\open{1}{y}{P}}$ $\tpcp$ z$\sim$B++x$\sim$C++$\Delta$P0
PER : Permutation $\Gamma$ ($\Delta$P++$\Delta$Q)
PER0 : Permutation ($\Delta$P++y$\sim$A)]) (x$\sim$B$\parr$C++$\Delta$P0)
\end{coq}\\ \hline
\begin{coq}
x(B).($\nu$ A.($\open{0}{1}{P}$ $\mid$ Q)) $\tpcp$ $\Gamma$.
\end{coq}
\end{tabular}
\\~\\

In order to prove the goal, $\Gamma$ must be replaced with the sublists typing
\coqe!P! and \coqe!Q!. Therefore, \coqe!PER0! needs to modified such that the
left-hand side is $\Delta$P and the right-hand side has \coqe!y$\sim$A!
removed. In other words, the binding of \coqe!y! within \coqe!$\Delta$P0! must
be extracted. Since many of the commuting conversion rules required this sort
of reasoning separating the work out into a tactic comes to mind. Indeed, Ltac
is well-suited for this type of work, searching the context and manipulating
hypotheses of a particular form. The Ltac script in Figure~\ref{fig:extract}
does the job\todo{keep (and explain) this?}.

@@

However, the task of removing the binding from the permutation is not
automated to keep information loss within the context transparent to an
individual proof. The tactic appears more general in this form and allows more
than one ordering of tactic sequences. feels like two jobs to remove it as
well as find it. although perhaps extract is not the right term if it doesn't
remove it

>>>>

\textbf{Weakening and Contraction.} Initially, an implicit notion of weakening
was chosen so as to be in line with \citeauthor{Wadler:2014}'s
definitions. However, explicit weakening is required to differentiate the
principal cut elimination rules for the interactions between servers and
clients; spawning and garbage collection (weakening). The paper presentation
does not specify where weakening could occur due to its implicit nature; it is
tacitly assumed weakening can always be pushed further up the derivation tree
to a base rule e.g. \coqe$cp_fwd$ rule. The same is true of contraction where
it is defined as a substitution of names on process terms, not a
syntax-directed rule.

@@

\textbf{Empty Choice.} The development provides a restricted form of the empty
choice rule. In contrast to the arbitrary context in
\citeauthor{Wadler:2014}'s presentation~\cite{Wadler:2014}, \coqe!cp_empcho!
may only have a single binding in the context. Adding an arbitrary environment
to the empty choice constructor would prohibit \coqe$in_env_fv$ since its
process no longer contains all the names of its environment. Not only would
weakening be inadmissible if \coqe$in_env_fv$ did not hold, but also axiom cut
reduction:
\begin{coq}
Lemma reduce_axcut:
  forall P A (w : atom) $\Gamma$
         (NFV: w $\notin$ fv_proc P)
         (WT: $\nu$ A.(w $\link$ 0 $\mid$ P) $\tpcp$ $\Gamma$),
    (open_proc P w) $\tpcp$ $\Gamma$.
\end{coq}

@@

\subsubsection{Equivalences and Conversions}

@@

Regarding commuting conversions, many require analysing permutations
containing singleton bindings. Consider the following the Coq encoding of the
commuting conversion for output (cf. Figure~\ref{fig:ccoutput}):
\begin{coq}
Lemma reduce_cc_multi_one:
  forall P Q R (x:atom) A B $\Gamma$ (LCQ: lc_proc Q)
         (WT: $\nu$ A.([B] x.(P $\mid$ Q) $\mid$ R) $\tpcp$ $\Gamma$),
    [B] x.($\nu$ A.($\permb{0}{1}{P}$ $\mid$ R) $\mid$ Q) $\tpcp$ $\Gamma$.
\end{coq}

Notice again, as in the case for \coqe!assoc_1!, it necessary to stipulate
that \coqe!Q! is locally closed to ensure \coqe!P! receives the channel
created by the outer cut as part of its environment.

After inverting \key{WT} and specialising the resulting hypotheses, regarding
the subprocesses, with fresh variables the proof state is as follows:

\begin{tabular}{l}
\begin{coq}
1 subgoal
(* ... other elements of the proof context ... *)
CPR : $\open{0}{y}{R}~\tpcp$ y$\sim$A++$\GD$Q
CPQ : Q $\tpcp$ x$\sim$C++$\GD$Q0
CPP : $\open{0}{z}{\open{1}{y}{P}}$ $\tpcp$ z$\sim$B++$\GD$P0
NIN : y $\notin$ {x} $\cup$ dom $\GD$Q0
PER : Permutation $\Gamma$ ($\GD$P++$\GD$Q)
PER0 : Permutation ($\GD$P++y$\sim$A) (x$\sim$B$\otimes$C++$\GD$P0++$\GD$Q0)
\end{coq}\\ \hline
\begin{coq}
x(B).($\nu$ A.($\open{0}{1}{P}$ $\mid$ Q)) $\tpcp$ $\Gamma$
\end{coq}
\end{tabular}
\\~\\

In order to prove the goal, $\Gamma$ must be replaced with the sublists typing
\coqe!P!, \coqe!Q! and \coqe!R!. Therefore, \coqe!PER0! needs to be modified
such that the left-hand side is $\GD$P and the right-hand side has
\coqe!y$\sim$A!  removed. In other words, the binding of \coqe!y! within
\coqe!$\GD$P0! must be extracted (by \coqe!NIN!, \coqe!y! must be in
\coqe!$\GD$P0!). Since many of the commuting conversion rules required this
sort of reasoning, separating the work out into a tactic appears
beneficial. Indeed, Ltac is well-suited for this type of work, searching the
context and manipulating hypotheses of a particular form. The Ltac script in
Figure~\ref{fig:extract} does the job and after executing
\coqe!extract_bnd y A! \key{PER0} becomes:
\begin{coq}
Permutation ($\GD$P++y$\sim$A) (x$\sim$B$\otimes$C++E1++y$\sim$A++E2++$\GD$Q0)
\end{coq}
where E1 and E2 are sublists of $\GD$P0.

@@

However, the task of removing the binding from the permutation is not
automated to keep information loss within the context transparent to an
individual proof. The tactic is more modular and general in this form,
permitting composition and permutation of tactic sequences.
